---
title: "R Notebook"
output: html_notebook
---


When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
```{r}
#data set
library("readxl")
library("dplyr")
library("xlsx")
library("tidyr")
data <- read_excel("C:\\Users\\antha\\Downloads\\uscensus.xlsx")
df <- data.frame(data)
invisible(df %>% filter_all(any_vars(!is.na(.))))
invisible(df %>% filter_all(any_vars(complete.cases(.))))
rename(df, "High School Graduation Rate" = "HG", "Gross State Product" = "NGSP", "Personal Income, By State" = "NPI", "Unemployment Rate" = "UR", "Violent Crime Rate" = "VCR")


```
#data cleaning
```{r}
df <- subset(df, select = -NGSP)
#irrelevant/unnecessary
```
#data statistics
```{r}
CrimeRateMean <- mean(df$VCR)
UnemploymentRateMean <- mean(df$UR)
HighSchoolGradRate <- mean(df$HG)
PersonalIncomeMean <- mean(df$NPI)
tab <- matrix(c(CrimeRateMean, UnemploymentRateMean, HighSchoolGradRate, PersonalIncomeMean), ncol=4, byrow = TRUE)

CrimeRateSTD <- sd(df$VCR)
UnemploymentRateSTD <- sd(df$UR)
HighSchoolGradSTD <- sd(df$HG)
PersonalIncomeSTD <- sd(df$NPI)
CrimeRateMode <- mode(df$VCR)
UnemploymentRateMode <- mode(df$UR)
HighSchoolGradMode <- mode(df$HG)
PersonalIncomeMode <- mode(df$NPI)
CrimeRateMedian <- median(df$VCR)
UnemploymentRateMedian <- median(df$UR)
HighSchoolGradMedian <- median(df$HG)
PersonalIncomeMedian <- median(df$NPI)
tab <- rbind(tab, c(CrimeRateSTD, UnemploymentRateSTD, HighSchoolGradSTD, PersonalIncomeSTD))
tab <- rbind(tab, c(CrimeRateMode, UnemploymentRateMode, HighSchoolGradMode,PersonalIncomeMode))
tab <- rbind(tab, c(CrimeRateMedian, UnemploymentRateMedian, HighSchoolGradMedian, PersonalIncomeMedian))

colnames(tab) <- c('CrimeRate', 'UnemploymentRate', 'HighSchoolGraduation', 'PersonalIncome')
rownames(tab) <- c('United States Average', 'Standard Deviation', 'Mode', 'Median')
#mean, median, mode, standard deviation
```
#statistical tests
```{r}
df1 <- subset(df, select = -State)
cor(df1)
model <- lm(VCR ~ UR + NPI + HG, data = df1)
summary(model)
chisq <- lapply(df1[,-1], function(x) chisq.test(df1[,1], x)); chisq

#correlational model, chi-squared test, multiple linear regression

```
#data Visualization
```{r}
library(ggplot2)
ggplot(df1, aes(HG,NPI)) + geom_point() + geom_smooth(method = lm)
plot(df)
ggplot(df, aes(x=HG)) + geom_histogram(binwidth = 1.5) + geom_histogram(color = "black", fill = "red")
ggplot(df, aes(x=NPI)) + geom_histogram(binwidth = 1.5) + geom_histogram(color = "black", fill = "blue")
ggplot(df, aes(x=VCR)) + geom_histogram(binwidth = 1.5) + geom_histogram(color = "black", fill = "green")
ggplot(df, aes(x=UR)) + geom_histogram(binwidth = 1.5) + geom_histogram(color = "black", fill = "purple")

barplot(height = df$VCR,
        main = "Violent Crime Rate by State",
        xlab = "State",
        ylab = "Violent Crime Rate",
        names.arg = df$State,
        border = "black",
        col = "red",
        las = 2,
        cex.names = .43
        )
barplot(height = df$HG,
        main = "High School Graduation Rate by State",
        xlab = "State",
        ylab = "High School Graduation Rate",
        names.arg = df$State,
        border = "black",
        col = "red",
        las = 2,
        cex.names = .43
        )
barplot(height = df$NPI,
        main = "Personal Income by State",
        xlab = "State",
        ylab = "Pesonal Income",
        names.arg = df$State,
        border = "black",
        col = "red",
        las = 2,
        cex.names = .43
        )
barplot(height = df$UR,
        main = "Unemployment Rate by State",
        xlab = "State",
        ylab = "Unemployment Rate",
        names.arg = df$State,
        border = "black",
        col = "red",
        las = 2,
        cex.names = .43
        )
p <- 
  ggplot(df, aes(HG, UR , size = VCR, color=State)) +
  geom_point() +
  theme_bw() 

ggplotly(p)


#graphs showing correlation with line of best fit
#histograms showing skews for each variable

```




```{r}

```



```{r}
library(ggplot2)
library(plotly)

p <- 
  ggplot(df, aes(HG, UR , size = VCR, color=State)) +
  geom_point() + labs(x="High School Graduation Rate",  y = "Unemployment Rate", title= "Comparing GED, Violence, and Unemployment by State", subtitle = "Click or hover over a bubble for more details", caption = "Larger circles indicate higher violent crime rates.") +
  theme_bw() + scale_size(name="Violent Crime Rate by Size") + theme(legend.position = "right") 

p1 <- 
  ggplot(df, aes(NPI, VCR , size = HG, color=State)) +
  geom_point() + labs(x="Personal Income",  y = "Violent Crime Rate", title= "Comparing Income, Violence, and GED by State", subtitle = "Click or hover over a bubble for more details", caption = "Larger circles indicate higher GED completion rates.") +
  theme_bw() + scale_size(name="GED by Size") + theme(legend.position = "right")

ggplotly(p)
ggplotly(p1)
```

#building rShiny application
```{r}
library(shiny)
library(shinythemes)
library(dplyr)
library(readr)
library(ggplot2)
library(plotly)
library(readxl)
library(dplyr)
library(xlsx)
library(tidyr)
library(rsconnect)
data <- read_excel("C:\\Users\\antha\\Downloads\\uscensus.xlsx")
df <- data.frame(data)
invisible(df %>% filter_all(any_vars(!is.na(.))))
invisible(df %>% filter_all(any_vars(complete.cases(.))))
rename(df, "High School Graduation Rate" = "HG", "Gross State Product" = "NGSP", "Personal Income, By State" = "NPI", "Unemployment Rate" = "UR", "Violent Crime Rate" = "VCR")

ui <- fluidPage(
  titlePanel("Plots comparing several variables in different states"),
  mainPanel(
            plotlyOutput("ggp1"),
            plotlyOutput("ggp2")
            )  
  )


server <- function(input, output){
  output$ggp1 <- renderPlotly(ggplot(df, aes(NPI, VCR , size = HG, color=State)) +
  geom_point() + labs(x="Personal Income",  y = "Violent Crime Rate", title= "Comparing Income, Violence, and GED by State", subtitle = "Click or hover over a bubble for more details", caption = "Larger circles indicate higher GED completion rates.") +
  theme_bw() + scale_size(name="GED by Size") + theme(legend.position = "right"))
  
  output$ggp2 <- renderPlotly(ggplot(df, aes(HG, UR , size = VCR, color=State)) +
  geom_point() + labs(x="High School Graduation Rate",  y = "Unemployment Rate", title= "Comparing GED, Violence, and Unemployment by State", subtitle = "Click or hover over a bubble for more details", caption = "Larger circles indicate higher violent crime rates.") +
  theme_bw() + scale_size(name="Violent Crime Rate by Size") + theme(legend.position = "left"))
  


}

shinyApp(ui = ui, server = server)

rsconnect::deployApp()
 
```


```{r}
library(rsconnect)

rsconnect::setAccountInfo(name='amh371',                            token='6B493FEA8CC29BCEC10AE350E1F7A58F',
                          secret='TpQesRkLQqPwHEbDmQkDQKh0qaocHDajrsKGKSQb')

rsconnect::deployApp()

```

